---
title: "Tutorial: pull_data_synapse"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEncoding{UTF-8}
  %\VignetteIndexEntry{Tutorial: pull_data_synapse}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

# exit if user doesn't have synapser, a log in, or access to data.
if (genieBPC:::.is_connected_to_genie() == FALSE){
  knitr::knit_exit()
}
```

```{r, include=FALSE, eval=genieBPC:::.is_connected_to_genie()}
library(genieBPC)
library(dplyr)
library(tibble)
library(magrittr)
library(gt)

set_synapse_credentials()
```

## Introduction

The `pull_data_synapse()` function accesses the specified version of the clinical and genomic GENIE BPC data from [Synapse](https://www.synapse.org/#!Synapse:syn21226493/wiki/599164) and reads it into the R environment.
&nbsp;

This vignette will walk a user through the `pull_data_synapse()` function.


## Setup

Before beginning this tutorial, be sure to have a Synapse account. If you do not yet have a Synapse account, please follow the instructions below:

1. Register for a [Synapse account](https://www.synapse.org/#!Reigster0)
2. Use [this link](https://www.synapse.org/#!Team:3399797) to access the team list and request to join the team. Please include your full name and affiliation in the message before sending out the request.
3. Once the request is accepted, you may access the data in the [GENIE Biopharma collaborative projects](https://www.synapse.org/#!Synapse:syn21226493).

*Note: Please allow up to a week to review and grant access.*

&nbsp;

Users must log in to Synapse to access the data successfully. 

- To set your Synapse credentials during each session, call:

  `set_synapse_credentials(username = "your_username", password = "your_password")`
  
- To store authentication information in your environmental variables, add the following to your .Renviron file (tip: you can use `usethis::edit_r_environ()` to easily open/edit this file):

  `SYNAPSE_USERNAME = <your-username>`
  `SYNAPSE_PASSWORD = <your-password>`

Alternatively, you can pass your username and password to each individual data pull function if preferred, although it is recommended that you manage your passwords outside of your scripts for security purposes.


## Usage

Let's start by reviewing the `pull_data_synapse()` arguments.


```{r echo = FALSE, warning = FALSE, message = FALSE, eval=genieBPC:::.is_connected_to_genie()}
tibble::tribble(
  ~Argument,       ~Description,
  "`cohort`",       "Vector or list specifying the cohort(s) of interest. Must be one of 'NSCLC' (Non-Small Cell Lung Cancer), 'CRC' (Colorectal Cancer), or 'BrCa' (Breast Cancer)",
  "`version`",        "Vector specifying the version of the data. Must be one of the following: 'v1.1-consortium', 'v1.2-consortium', 'v2.1-consortium', 'v2.0-public'. When entering multiple cohorts, the order of the version numbers corresponds to the order that the cohorts are specified; the cohort and version number must be in the same order in order to pull the correct data.",
  "`download_location`",        "If `NULL` (default), data will be returned as a list of dataframes with requested data as list items. Otherwise, specify a folder path to have data automatically downloaded there. When a path is specified, data are not read into the R environment.",
  "`username`",        "Synapse username",
  "`password`",        "Synapse password"
) %>%
  gt::gt() %>%
  gt::fmt_markdown(columns = c(Argument)) %>%
  gt::tab_options(table.font.size = 'small',
                  data_row.padding = gt::px(1),
                  summary_row.padding = gt::px(1),
                  grand_summary_row.padding = gt::px(1),
                  footnotes.padding = gt::px(1),
                  source_notes.padding = gt::px(1),
                  row_group.padding = gt::px(1))
```

&nbsp;

## Examples


### Example 1

Pull version 2.1-consortium of the NSCLC data from Synapse and store in the local environment.

```{r, eval = FALSE}
nsclc_2_1 = pull_data_synapse("NSCLC", version = "v2.1-consortium")
```


The resulting `nsclc_data` object is a list of elements, such that each element represents a dataset:

```{r, include = FALSE, eval=genieBPC:::.is_connected_to_genie()}
nsclc_2_1 = pull_data_synapse("NSCLC", version = "v2.1-consortium")
```


```{r, echo = FALSE, message = FALSE, eval=genieBPC:::.is_connected_to_genie()}
ls(nsclc_2_1$NSCLC_v2.1) %>% 
  as.data.frame() %>% 
  mutate(clinical_vs_genomic = case_when(. %in% c("cna", "fusions",
                                                  "mutations_extended") ~ "Genomic Data",
                                         TRUE ~ "Clinical Data")) %>% 
  gt::gt() %>% 
  gt::cols_label(.data = .,
                 `.` = md("**Datasets Returned**"),
             clinical_vs_genomic = md("**Clinical/Genomic Data**"))
```



### Example 2

Pull version 2.1-consortium of the NSCLC data and version 1.1-consortium of the CRC data.

```{r, eval = FALSE}
pull_data_synapse(c("NSCLC", "CRC"), 
                  version = c("v2.1-consortium","v1.1-consortium"))
```



### Example 3

Pull version 1.1-consortium of the NSCLC data and version 1.1-consortium of the CRC data.

```{r, eval = FALSE}
pull_data_synapse(c("NSCLC", "CRC"),
                  version = c("v1.1-consortium", "v1.1-consortium"))
```



## Future Work

* As of June 2022, the non-small cell lung cancer (NSCLC), colorectal cancer (CRC), and breast cancer (BrCa) cohorts are available to access. Future cohorts will include pancreas, prostate and bladder cancers.
